---
title: "EDA of Seattle Building Energy Benchmarks"
author: "Christopher Lacrampe"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:

    code_folding: show
    highlight: tango
    theme: yeti
    toc: yes
    toc_depth: 4
    toc_float: yes
---

# 1: Exploring the structure of the datat

#### Loading the dataset and libraries

```{r}
energyDat2016 = read.csv("sea-building-energy-benchmarking/2015-building-energy-benchmarking.csv")
library(gmodels)
library(tidyverse)
```

#### Asessing the structure of the data

```{r}
str(energyDat2016)
```

#### Explaining the features of the data:

For more details concerning the variables, please see the data dictionary stored in "variable-description.txt" within the sea-building-energy-benchmarking folder:

# 2: Exploring the numerical variables
We'll focus on the variables: **YearBuilt**, **NumberOfFloors**, **PropertyGFATotal** (building and parking gross floor area), **ENERGYSTARScore**, **SiteEUIWN** (kBtu/sf, normalized for weather), **Electricity** (kBtu), **NaturalGas** (kBtu), **TotalGHGEmissions** (kilograms of c02 equivalent per square foot)

#### Exploring the central and spread measures for numeric features

```{r}
summary(energyDat2016$YearBuilt)
summary(energyDat2016$NumberofFloors)
summary(energyDat2016$PropertyGFATotal)
summary(energyDat2016$ENERGYSTARScore)
summary(energyDat2016$SiteEUIWN.kBtu.sf.)
summary(energyDat2016$Electricity.kBtu.)
summary(energyDat2016$NaturalGas.kBtu.)
summary(energyDat2016$GHGEmissionsIntensity.kgCO2e.ft2.)
```

From these descriptive statistics we can see that for:  
**YearBuilt** we have a range of 115 years, with a left-skewed distribution  
**NumberofFloors** we have a very large range from 0 to 99 with a median of 4 and a mean of 4.682 (right-skewed). We also have 8 unreported values  
**PropertyGFATotal** we have a range from 20000 to 2200000 with a median of 43855 and a mean of 90773 (the data is very right-skewed). We may have to adjust for some outliers  
**ENERGYSTARScore** has a range from 1 to 100, with a median of 76 and a mean of 67.89 (left-skewed). We have the greatest number of missing values at 780.
**SiteEUIWN.kBtu.sf.** has a range from 0 to 800.60, with a median of 40.9 and a mean of 57.58 (again the data is right-skewed)  
**Electricity.kBtu.** has a range from 0 to 4726322 with a median of 1160236 and a mean of 3473209 (the data is also very right-skewed). Also important to note are the 10 unreported values  
**NaturalGas.kBtu.** has a range of 0 to 6448438 with a median of 294577 and a mean of 125330 (the data is also very right-skewed). This variable is also missing ten values--perhaps from the same observations as the missing values for Electricity.kBtu  
**GHGEmissionsIntensity.gjC02e.ft2.** has a range from 0 to 31 with a median of 0.46 and a mean of 0.9853 (the data is also right skewed). This variable is also missing ten values as well  

#### Creating boxplots and histograms for the numerical variables  

```{r}
boxplot(energyDat2016$YearBuilt, main = "Year Built", ylab = "Year")
hist(energyDat2016$YearBuilt, main = "Histogram of the Year Built", xlab = "Year")
boxplot(energyDat2016$PropertyGFATotal, main = "Boxplot of Building and Parking Gross Floor area", ylab = "Total square feet")
hist(energyDat2016$PropertyGFATotal, main = "Histogram of Building and Parking Gross Floor area", xlab = "Total square feet")
boxplot(energyDat2016$ENERGYSTARScore, main = "Boxplot of ENERGYSTARScore", ylab = "Score out of 100")
hist(energyDat2016$ENERGYSTARScore, main = "Histogram of ENERGYSTARScore", xlab = "Score out of 100")
boxplot(energyDat2016$Electricity.kBtu., main = "Boxplot of Electricity usage", ylab = "kilo Brittish thermal units")
hist(energyDat2016$Electricity.kBtu., main = "Histogram of Electricity usage", xlab = "kilo Brittish thermal units")
boxplot(energyDat2016$NaturalGas.kBtu., main = "Boxplot of Natural gas usage", ylab = "kilo Brittish thermal units")
hist(energyDat2016$NaturalGas.kBtu., main = "Histogram of Natural gas usage", xlab = "kilo Brittish thermal units")
```

#### Computing variance and standard deviations for the numerical variables

```{r}
var(energyDat2016$YearBuilt, na.rm=TRUE)
sd(energyDat2016$YearBuilt,na.rm=TRUE)
var(energyDat2016$NumberofFloors,na.rm=TRUE)
sd(energyDat2016$NumberofFloors,na.rm=TRUE)
var(energyDat2016$PropertyGFATotal,na.rm=TRUE)
sd(energyDat2016$PropertyGFATotal,na.rm=TRUE)
var(energyDat2016$ENERGYSTARScore,na.rm=TRUE)
sd(energyDat2016$ENERGYSTARScore,na.rm=TRUE)
var(energyDat2016$SiteEUIWN.kBtu.sf.,na.rm=TRUE)
sd(energyDat2016$SiteEUIWN.kBtu.sf.,na.rm=TRUE)
var(energyDat2016$Electricity.kBtu.,na.rm=TRUE)
sd(energyDat2016$Electricity.kBtu.,na.rm=TRUE)
var(energyDat2016$NaturalGas.kBtu.,na.rm=TRUE)
sd(energyDat2016$NaturalGas.kBtu.,na.rm=TRUE)
var(energyDat2016$GHGEmissionsIntensity.kgCO2e.ft2.,na.rm=TRUE)
sd(energyDat2016$GHGEmissionsIntensity.kgCO2e.ft2.,na.rm=TRUE)
```

# 3: Exploring the categorical variables  
Let's take a deeper look into the categorical variables of interest: **NeighborhoodProperty**, **LargestPropertyUseType**, and **Outlier**  
```{r}
table(energyDat2016$Neighborhood)
table(energyDat2016$LargestPropertyUseType)
table(energyDat2016$Outlier)
```

#### Finding the mode/central measure of the categorical variables
```{r}
names(sort(-table(energyDat2016$Neighborhood)))[1]
names(sort(-table(energyDat2016$LargestPropertyUseType)))[1]
names(sort(-table(energyDat2016$Outlier)))[1]
```

#### Finding proprotion tables 

```{r}
round(prop.table(table(energyDat2016$Neighborhood))*100, digits = 3)
round(prop.table(table(energyDat2016$LargestPropertyUseType))*100, digits = 3)
round(prop.table(table(energyDat2016$Outlier))*100, digits = 3)
```

# 4: Exploring relationships between two variables