---
title: "EDA of Seattle Building Energy Benchmarks"
author: "Christopher Lacrampe"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:

    code_folding: show
    highlight: tango
    theme: yeti
    toc: yes
    toc_depth: 4
    toc_float: yes
---

# Exploring the structure of the datat

#### Loading the dataset and libraries

```{r}
energyDat = read.csv("sea-building-energy-benchmarking/2015-building-energy-benchmarking.csv")
library(gmodels)
library(tidyverse)
```

#### Asessing the structure of the data

```{r}
str(energyDat)
```

#### Explaining the features of the data:

For more details concerning the variables, please see the data dictionary stored in "variable-description.txt" within the sea-building-energy-benchmarking folder:

# Exploring the numerical variables
We'll focus on the variables: **YearBuilt**, **NumberOfFloors**, **PropertyGFATotal** (building and parking gross floor area), **ENERGYSTARScore**, **SiteEUIWN(kBtu/sf, normalized for weather)**, **Electricity** (kBtu), **NaturalGas** (kBtu), **TotalGHGEmissions**(kilograms of c02 equivalent per square foot)

```{r}
summary(energyDat$YearBuilt)
summary(energyDat$NumberofFloors)
summary(energyDat$PropertyGFATotal)
summary(energyDat$SiteEUIWN.kBtu.sf.)
summary(energyDat$Electricity.kBtu.)
summary(energyDat$NaturalGas.kBtu.)
summary(energyDat$GHGEmissionsIntensity.kgCO2e.ft2.)
```

