---
title: "Regression-SEA-building-energy-bechmarking"
author: "Christopher Lacrampe"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:

    code_folding: show
    highlight: tango
    theme: yeti
    toc: yes
    toc_depth: 5
    toc_float: yes
---

The aim of this predictive model is to be able to identify buildings on the basis of being able to obtain Energy Star Certification (Energy Star Score >= 75). By training a Regression, we can measure the factors that impact energy star score the most, which can help buildings in the Seattle area that have not enrolled in the energy star program, but have access to the same variables as our training and test sets (so that they can improve their score in the future).

# 1. Loading, cleaning, and preparing the data/libraries

### 1.1 Loading libraries and datasets

```{r}
library(psych)
library(tidyverse) # has useful libraries for cleaning and filtering data
energyDat2016 = read.csv("sea-building-energy-benchmarking/2015-building-energy-benchmarking.csv", stringsAsFactors = FALSE)
```

### 1.2 Cleaning the data

#### 1.2.1 Removing outliers
Our dataset has come prepared with an **Outlier** feature where the data controller has flagged outliers as "High" and "Low", we will remove those outliers

```{r}
energyDat2016 = tbl_df(filter(energyDat2016, Outlier == ""))
energyDat2016 = as.data.frame(energyDat2016)
```

#### 1.2.2 Create a dataframe utilizing the variables we want to include in our model (this strips redundant variables that are identical to the others just measured in different units)

```{r}
modelDat = select(energyDat2016, ENERGYSTARScore, YearBuilt, PropertyGFATotal, SiteEUIWN.kBtu.sf., Electricity.kBtu., NaturalGas.kBtu., GHGEmissionsIntensity.kgCO2e.ft2.)
# remove NA values
modelDat = na.omit(modelDat)
```

# 2. Exploring the data

#### 2.1 Enhanced scatterplot matrix

```{r}
pairs.panels(modelDat)
```

In the upper right trianlge of the matrix we see the correlations between each of the numerical features. Along the main diagonal we see a histogram of each feature's distribution. In the lower triangular portion we see scatterplots between two features. The red line is a loess curve that somewhat models the relationship between the two variables. The elipse represents the strength of the correlation with the center dot indicating the mean point (mainy of the elipses are centered, so the correlation between features isn't very strong).

# 3. Training a model on the data

### 3.1 Use the lm() method to fit a linear model to the data. Name it energy_model

```{r}
energy_model = lm(ENERGYSTARScore ~ YearBuilt + PropertyGFATotal + SiteEUIWN.kBtu.sf. + Electricity.kBtu. + NaturalGas.kBtu. + GHGEmissionsIntensity.kgCO2e.ft2., data = modelDat)
```

### 3.2 Show the estimated Beta coefficients. Explain the results

```{r}
summary(energy_model)
```

From this summary output we can see that for each increase in YearBuilt we see a decrease in ENERGYSTARSCore of -5.197e-04. As the PropertyGFATotal increases by one square foot, we can expect an increase in ENERGYSTARScore of 4.047e-06. As Electricity.kBtu. increasees by one kBtu the ENERGYSTARScore increases 3.777e-07. As NaturalGas.kBtu. increases by one kBtu, ENERGYSTARScore decreases by -2.796e-07. As GHGEmissionsIntensity.kgCO2e.ft2. increases by one unit, ENERGYSTARScore increases by 2.325e points. The Adjusted R-squared of 0.1471 indicates that 14.71% of the variation in ENERGYSTARScore is explained by the model.

# 4. Improving model performance

### 4.1

